---
title: 'Take_home_Stats_test '
author: "Nana  Boateng"
date: '`r Sys.Date()`'
Time: '`r Sys.time()`'
output: html_document
  #pdf_document: default
---

```{r setup, include=FALSE}
list=c("tidyverse","stringr","forcats","ggmap","rvest","tm","SnowballC","dplyr","calibrate","doParallel",
       "stringi","ggplot2","maps","httr","rsdmx","devtools","plyr","dplyr","ggplot2","caret","elasticnet",
       "magrittr","broom","glmnet","Hmisc",'knitr',"RSQLite","RANN","lubridate","ggvis","plotly","lars",
       "ggcorrplot","GGally","ROCR","lattice","car","corrgram","ggcorrplot","parallel","readxl","ggmosaic",
       "vcd","Amelia","d3heatmap","ResourceSelection","ROCR","plotROC","DT","aod","mice","Hmisc","data.table","pls","plsdepot","gvlma","ggthemes","ggfortify","grid","ggpmisc")



list_packages <- list
new.packages <- list_packages[!(list_packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

R<-suppressWarnings(suppressMessages(sapply(list, require, character.only = TRUE)))

setwd("/Users/nanaakwasiabayieboateng/Documents/memphisclassesbooks/DataMiningscience/UnitedAirlines")

```


```{r}
library(readxl)
# read_excel reads both xls and xlsx files
uniteddata=read_excel("Take_home_Stats_test_Data.xlsx")
DT::datatable(uniteddata)

```




##**Model One**
For a unit change in X1 keeping X2 constant, the mean of Y increases by 2.149586 and 
unit change in X2 keeping X1 constant, the mean of Y increases by 3.465170

```{r}
model1=glm(Y~X1+X2,data=uniteddata)

broom::tidy(model1)
```

```{r}

head(broom::augment(model1))
```

```{r}





broom::glance(model1)
```

##**Exploratory Data Analysis**
```{r}
Y=uniteddata$Y
X2=uniteddata$X2
#formula <- y ~ poly(x, 3, raw = TRUE)
formula <- Y ~ X2
#formula <- y ~ poly(x, raw = TRUE)
fill <- "#4271AE"
line <- "#1F3552"

#p <-
  ggplot(uniteddata, aes(x=X1, y=Y)) + geom_point(shape=1) + geom_smooth(method=lm, se=FALSE) +
      ggtitle(" regression line") +
      scale_x_continuous(name = "X1") +
      scale_y_continuous(name = "Y") +
theme_economist() +
      theme(axis.line.x = element_line(size=.5, colour = "black"),
            axis.line.y = element_line(size=.5, colour = "black"),
            axis.text.x=element_text(colour="black", size = 9),
            axis.text.y=element_text(colour="black", size = 9),
            panel.grid.major = element_line(colour = "#d3d3d3"),
            panel.grid.minor = element_blank(),
            panel.border = element_blank(), panel.background = element_blank(),
            plot.title = element_text(family = "Tahoma"),
            text=element_text(family="Tahoma"))+
  stat_poly_eq(aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
                   formula = formula, rr.digits = 3, coef.digits = 2, parse = TRUE)
#ggplotly(p)

```



```{r}
Y=uniteddata$Y
X2=uniteddata$X2
formula <- Y ~ X2
fill <- "#4271AE"
line <- "#1F3552"

ggplot(uniteddata, aes(x=X2, y=Y)) + geom_point(shape=1) + geom_smooth(method=lm, se=FALSE) +
      ggtitle(" regression line") +
      scale_x_continuous(name = "X1") +
      scale_y_continuous(name = "Y") +
      theme(axis.line.x = element_line(size=.5, colour = "black"),
            axis.line.y = element_line(size=.5, colour = "black"),
            axis.text.x=element_text(colour="black", size = 9),
            axis.text.y=element_text(colour="black", size = 9),
            panel.grid.major = element_line(colour = "#d3d3d3"),
            panel.grid.minor = element_blank(),
            panel.border = element_blank(), panel.background = element_blank(),
            plot.title = element_text(family = "Tahoma"),
            text=element_text(family="Tahoma"))+
   stat_poly_eq(formula = y ~ x, eq.with.lhs=FALSE,
      aes(label = paste("hat(italic(y))","~`=`~",..eq.label..,"~~~", ..rr.label.., sep = "")),parse = TRUE)

```



```{r}

fill <- "#4271AE"
line <- "#1F3552"

p <-ggplot(uniteddata, aes(x=X1, y=Y)) + geom_point(shape=1) + geom_smooth(method=lm, se=FALSE) +
      ggtitle(" regression line") +
      scale_x_continuous(name = "X1") +
      scale_y_continuous(name = "Y") +
theme_economist() +
      theme(axis.line.x = element_line(size=.5, colour = "black"),
            axis.line.y = element_line(size=.5, colour = "black"),
            axis.text.x=element_text(colour="black", size = 9),
            axis.text.y=element_text(colour="black", size = 9),
            panel.grid.major = element_line(colour = "#d3d3d3"),
            panel.grid.minor = element_blank(),
            panel.border = element_blank(), panel.background = element_blank(),
            plot.title = element_text(family = "Tahoma"),
            text=element_text(family="Tahoma"))
ggplotly(p)

```



**Checking Regression Assumptions**
1) The mean of the residual errors is close to 0.

```{r}

mean(model1$residuals)
```
```{r}
#summary(model1)
#model1$residuals
#model1$fitted
#dataplot=data_frame(residuals=model1$residuals,fitted=model1$fitted,stdred=model1$std.resid)
par(mfrow = c(1, 2))
library(ggfortify)

ggplot2::autoplot(model1, which = 1:6, ncol = 3, label.size = F)+ theme_bw()

```






```{r}
library(ggfortify)
myfortdata = fortify(USArrests)
myfortdata%>%head()%>%DT::datatable()
#head(myfortdata)%>%head()%>%DT::datatable()

dataplot=data_frame(residuals=model1$residuals,fitted=model1$fitted)

p1<-ggplot(dataplot, aes(fitted, residuals))+geom_point()+
    geom_smooth(method="loess")+geom_hline(yintercept=0, col="red", linetype="dashed")+
    xlab("Fitted values")+ylab("Residuals")+
    ggtitle("Residual vs Fitted Plot")+theme_bw()+geom_point(shape=1,color="purple")
ggplotly(p1)
```
```{r}
#===========================================================================
# 
# broom augment and cbind.data.frame  data add regression diagnostics to data
#===========================================================================

par(mfrow=c(2,2))
myudata=cbind.data.frame(uniteddata,broom::augment(model1))
p=ggplot(data = myudata, aes(x = .fitted, y = .resid)) +
  geom_hline(yintercept = 0, colour = "firebrick3") +geom_smooth(method="loess")+
 geom_point()+xlab("Fitted values")+ylab("Residuals")+ggtitle("Residual vs Fitted Plot")+theme_bw()
ggplotly(p)

#===========================================================================
# 
# fortify data add regression diagnostics to data
#===========================================================================

myundata= fortify(model1)

pp=ggplot(data = myudata, aes(x = .fitted, y = .resid)) +
  geom_hline(yintercept = 0, colour = "firebrick3") +
  geom_point() +
  geom_smooth(se = FALSE)+theme_bw()+
  xlab("Fitted Values")+ylab("Residuals")+
  ggtitle("Homoscedasticity assumption")
ggplotly(pp)

```

```{r}
myundata%>%head()
```

```{r}
p2=ggplot(data = myundata, aes(sample = .stdresid)) +
  stat_qq() +geom_abline(colour = "firebrick3")+
    ggtitle("Theoritcal Quantiles/Normality assumption")+theme_bw()
ggplotly(p2)
```



```{r}
p3=ggplot(data = myundata, aes(x = .fitted, y = .stdresid)) +
  geom_hline(yintercept = 0, colour = "firebrick3") +
  geom_point()+ geom_smooth(se=T,method = 'loess')+theme_bw()+
  ggtitle("Standadized Residuals vs Fitted")+
  xlab("Fitted Value")+
ylab((("Standardized residuals|")))
ggplotly(p3)
```




```{r}
p4=ggplot(data = myundata, aes(x = .hat, y = .stdresid)) +
  geom_point() +
  geom_smooth(se = FALSE,method = 'loess')+ geom_smooth(se=T,method = 'loess')+theme_bw()+
  ggtitle("Residuals vs. leverages")+
  xlab("Fitted Value")+
ylab("Standardized residuals|")
ggplotly(p4) 
```








```{r}
# Points size reflecting Cook's distance
p5=ggplot(data = myundata, aes(x = .fitted, y = .resid, size = .cooksd)) +
  geom_hline(yintercept = 0, colour = "firebrick3") +
  geom_point() +scale_size_area("Cook’s distance")+
  ggtitle("Points size reflecting Cook's distance")+
  theme_bw()+xlab("Fitted Values")+ylab("Residuals")
ggplotly(p5) 
```






```{r}
# Residuals vs. leverages with observation number
p6=ggplot(data = myundata, aes(x = .hat, y = .stdresid)) +
  geom_point() +
  geom_smooth(se = FALSE,method = 'loess') +
  geom_text(label = rownames(myundata), size = 4) +
  geom_hline(yintercept = 2, lty = "dotted", colour = "slateblue1") +
  geom_hline(yintercept = -2, lty = "dotted", colour = "slateblue1")+
  theme_bw()+xlab("Fitted Values")+ylab("Standidized Residuals")+theme_bw()
ggplotly(p6) 
```



```{r}
p7=ggplot(data = myundata, aes(x =X1, y = Y)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), colour = "red", se = FALSE) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2) + I(x^3), colour = "blue", se = FALSE)+
  theme_bw()
ggplotly(p7)
```




```{r}


#we define our own function mean of log
#and change geom to line
meanlog <- function(y) {mean(log(y))}
ggplot(myundata, aes(x=X1, y=Y)) + 
  stat_summary(fun.y="meanlog", geom="line")
```



```{r}
tp <- ggplot(myundata, aes(Y))+geom_density()
tp

```






2) Homoscedasticity of residuals or equal variance. This assumption is satisfied by 
normality of the residuals from the plots below.


3)No autocorrelation of residuals present from the plot. the second line in the plot is close to 0.
```{r}
# Method 1: Visualise with acf plot

acf(model1$residuals)

```
```{r}
# Method 3: Durbin-Watson test
lmtest::dwtest(model1)
```

5)The X variables and residuals are uncorrelated
```{r}

cor.test(uniteddata$X1, model1$residuals) 
cor.test(uniteddata$X2, model1$residuals) 
```

5)The variability in X values is positive.The X variables in the sample must all not 
be the same or nearly the same.


```{r}
var(uniteddata$X1)
var(uniteddata$X2)
```


5) No perfect multicollinearity
There is no perfect linear relationship between explanatory variables.The varianc einflation factor values is small(<10) for both explanatory variables signifying no problem with multicollinearity.
```{r}
vif(model1)
```

The correlation between X1 and X2 is very high (0.7111551). This  vilates the linear regression assumption.
```{r}
library(corrplot)
corrplot(cor(uniteddata[, -c(1,2)]))
```

```{r}
cor(uniteddata[, -c(1,2)])
```

Check assumptions automatically
```{r}
#par(mfrow=c(2,2))  # draw 4 plots in same window
#library(gvlma)

model=lm(Y~X1+X2,data=uniteddata)

gvlma::gvlma(model)



```


```{r}
#DT:datatable(influence.measures(mod))
#DT::datatable(data.table(influence.measures(mod)))
#data.table(influence.measures(mod))
DT::datatable((influence.measures(model)$infmat))
```






##**Model Two**
Principal components regression/Analysis or varoius forms of Factor Analysis can be used when the explanatory variables are correlated. 
```{r}


model2=pls::plsr(Y~X1+X2,data=uniteddata,ncomp=1)

summary(model2)



```


```{r}
3
#biplot(model2)
#plot(model2, plottype = "biplot")
```


```{r}
library(plsdepot)
model3=plsdepot::plsreg1(uniteddata[,3:4],uniteddata["Y"],comps=1, crosval=TRUE)


plot(model3)
```



The chi-square statistic and p-value in factanal are testing the hypothesis that the model fits the data perfectly.
About 77.3% of the variance explained by one factors is very high.
```{r}
factanal(uniteddata[,-1],1)
```



##**Model Two**
The data here was restructured into three columns which are the number of trials(1,..13), Driving school(Yes/No ) and the Frequency of the number of Trials. 
```{r}
uniteddata2=data_frame(No.Trial=c(seq(1:13)),No=c(29,16,17,4,3,9,4,5,1,1,1,3,7)
,Yes=c(198,107,55,38,18,22,7,9,5,3,6,6,12))
 

m1=reshape2::melt(uniteddata2,id.vars ="No.Trial",variable.name ="Driving.school",value.name ="Frequency")                        
m2= reshape2::dcast(reshape2::melt(uniteddata2,id.vars ="No.Trial"),No.Trial~variable )

DT::datatable(m1)
```
Chisquare Test.
```{r}
xtabs(No.Trial~Driving.school+Frequency, data=m1)

```

```{r}
chisq.test(xtabs(No.Trial~Driving.school+Frequency, data=m1),simulate.p.value = T)%>%tidy()

```

The frequency of those passing the driving test who attended a driving school is higher than 
those who didn't attend a driving school. Attending a driving school increases your probability of passing a driving test.
```{r}
p=ggplot(m1, aes(No.Trial,Frequency, color=Driving.school)) + 
  geom_line() + 
  geom_point()+theme_bw()+xlab("Number of Trials")
ggplotly(p)
```


a) Distribution plot of the Frequency of Number of trials for both those who had training from 
driving school and those who didn't. The distribution is skewed to the right. A logarithmic 
transformation could improve the skewness of the distribution.
```{r}
pq1=qplot( m1$Frequency, geom="histogram",binwidth = 2)+theme_minimal()+xlab("Number of Trials")
ggplotly(pq1)
```

Taking log of the Frequency of Number of trials for both those who had training from 
driving school and those who didn't. The regular assumption on the explanatory variables is that of normality.Even with the logarithmic transformation, the distribution doest appear to be normal. 
```{r}
pq=qplot(log( m1$Frequency), geom="histogram",binwidth = 0.5)+theme_minimal()+xlab("log(Number of Trials)")
ggplotly(pq)
```


c)

For a unit increase in the number of students  attending a driving-school driver to pass the test in a trial,the mean Frequency of passing thedriving test increases by 1.90802228 compared to those who did not attend driving school.
 
```{r}
model4=glm(log(Frequency)~(No.Trial)+Driving.school+No.Trial*Driving.school,data=m1)
#summary(model4)

broom::tidy(model4)




```
```{r}

head(broom::augment(model4))


```

```{r}
broom::glance(model4)
```

b)
A logistic regression of whether a person attended driving school or not on the number of trials and frequency  of attempts in passing the driving test does  have  significant variables at a level of  0.05. The frequency of passing a test and the number of trials are significant at 0.05 level.
```{r}
index <- createDataPartition(m1$Driving.school,p=0.70, list=FALSE)

trainSet <- m1[index,]

testSet <- m1[-index,]



model5=glm(Driving.school~No.Trial+log(Frequency),data=trainSet,family="binomial")



knitr::kable(confint(model5))

```

```{r}
tidy(model5)
```

For a unit increase in the frequency of a passing a dricing test, log odds of passing a driving test when one has taking a driving test test increases by about 200% compared to those who did not take the driving test.
```{r}
coef(model5)
```


The odds of passing a driving test when one has taking a driving test test increases by about 800% compared to those who did not take the driving test
```{r}




## odds ratios only)

## odds ratios and 95% CI
exp(cbind(OR = coef(model5), confint(model5)))


```

```{r}
# Predict using the test data
#testSet 
pred<-predict(model5,testSet)


# 
# # Print, plot variable importance
 print(varImp(model5, scale = FALSE))
# 
# plot(varImp(model5, scale = FALSE), main="Variable Importance using logistic/glm")
# 
# 
# confusionMatrix(testSet$Hypertension,pred)

```
```{r}
 my_data=data.frame(cbind(predicted=pred,observed=testSet$Driving.school))
 
ggplot(my_data,aes(predicted,observed))+geom_point()+geom_smooth(method=lm)+ggtitle('logistic model')
```




```{r}
predict <- predict(model5,testSet, type = 'response')

pp=if_else(predict >0.5,"Yes","No")

tab=table(testSet$Driving.school, pp)
tab
```

```{r}
predict <- predict(model5,testSet, type = 'response')

predict=if_else(predict >0.5,"Yes","No")
#predict
data.frame(observed=testSet$Driving.school,predicted=predict)
#cbind(m1$Driving.school, predict)

```





```{r}
#==================================================================
#ROCR Curve
#==================================================================




p <- ggplot(data.frame(trainSet), aes(d = as.numeric(Driving.school), m = Frequency)) + geom_roc()+ style_roc()


plot_interactive_roc(p)

```


The probability of   be the probability of a no-driving-school driver to pass the test in a trial  ,Pn is 0.1706485  and Ps be the probability of a driving-school driver to pass the test in a trial is 0.8293515. This is nearly 5 times the rate for those not attending driving school.  
```{r}
tapply(m1$Frequency,m1$Driving.school, sum)/sum(m1$Frequency)
```


